---
title: "2016 Burnham & Lehman Queen Experirment"
author: "P. Alexander Burnham"
date: "September 14, 2016"
output: pdf_document
---

**Metadata:** 

**Author:** P. Alexander Burnham

**Date:** 10 August 2016

**Data Set:** These data were collected in 2016 in New York state by Andre Burnham and the Hamilton College bee research team from two yards in NY owned by one beekeeper who volunteered for the study.

**Data Source:** 2016 Hamilton Bee Research Project

**Funding Source:** Hamilton College undergraduate research grants

**Data Collection:** Weights of colonies were obtained by summing the weights of individual hive bodies and supers for each colony at each yard (add more about protocols for other assays).

**Columns:** (from left to right) Field ID: Id including number location and treatment, Mass (biomass minus supers in lbs) Time (T1-T4), Origin (California and Local), Yard (1 or 2),Nosema (spores per bee) only two time steps (T1 and T3),Varroa (mites per 300 bees),Brood (frames of brood in colony),NosemaPA (binary data for nosema),VarroaPA (binary data for mites),
Rows: Data points for all columns in order from each collection event

**Missing values:** NA


```{r}

#Preliminaries:

ls()
rm(list=ls())

# Set Working Directory 
setwd("~/Desktop/QueenExperimentBurnham")

# read in data for all three .csv files:
QueensDF <- read.table("2016QueensHam.csv",header=TRUE,sep=",",stringsAsFactors=FALSE)

VirusDF <- read.table("RNAVirus.csv",header=TRUE,sep=",",stringsAsFactors=FALSE)

PollenDF <- read.table("PollenQueens.csv",header=TRUE,sep=",",stringsAsFactors=FALSE)

```



```{r}

# looking at colony mass through time

#subsetting by removing rows that have missing data:
QueensMassDF <- QueensDF[-c(95,98,102,104,105),]

# create summary and sd and se using plyr
library(plyr)
MassSummary <- ddply(QueensMassDF, c("Origin", "Time"), summarise, 
                  n = length(Mass),
                  mean = mean(Mass),
                  sd = sd(Mass),
                  se = sd / sqrt(n))

print(MassSummary)

```



```{r}
library(ggplot2)

# remove generic time steps and add the correct days for correct time scale
x <- MassSummary[,-2]
times <- c(1,24,43,57,1,24,43,57)
x <- data.frame(times, x)

#Create plot in ggplot 
plot <- ggplot(data = x, 
               aes(x = times, 
                   y = mean, 
                   group = Origin, 
                   colour = Origin)
               ) + geom_line(size=1.5) + geom_point(size=4) + scale_colour_manual(values = c("dodgerblue4", "black")) + labs(x = "Time (days after requeening)", y = "Colony Mass (lbs)") + coord_cartesian(ylim = c(20, 70)) + geom_errorbar(aes(ymin = mean - se, ymax = mean + se, width = 0.9)) 

# add a theme and add asterix for significance 
plot + scale_fill_brewer(palette = "Paired") + theme_minimal(base_size = 17) + annotate(geom = "text", x = 24, y = 44, label = "*",cex = 8) + annotate(geom = "text", x = 57, y = 66, label = "*",cex = 8) 

```




```{r}

#Check for normality and log transform
#hist(QueensMassDF$Mass)
#hist(log(QueensMassDF$Mass))

#Subset data into 4 time steps (T1-T4)
QMassT1 <- QueensMassDF[c(1:40),]
QMassT2 <- QueensMassDF[c(41:75),]
QMassT3 <- QueensMassDF[c(76:106),]
QMassT4 <- QueensMassDF[c(107:132),]

# run 4 t-tests comparing california to local colony mass
ComparingQueensMassT1 <- aov(log(Mass)~Origin, data=QMassT1)
ComparingQueensMassT2 <- aov(log(Mass)~Origin, data=QMassT2)
ComparingQueensMassT3 <- aov(log(Mass)~Origin, data=QMassT3)
ComparingQueensMassT4 <- aov(log(Mass)~Origin, data=QMassT4)

# view the summary of the stats:
summary(ComparingQueensMassT1)
summary(ComparingQueensMassT2)
summary(ComparingQueensMassT3)
summary(ComparingQueensMassT4)

```

```{r}

#subsetting by removing rows that have missing data:
QueensNosemaDF <- QueensDF[-c(35,(41:75)),]

times <- c(rep("T1",39),rep("T2",31),rep("T3",26))

QueensNosemaDF <- data.frame(QueensNosemaDF, times)

# create summary and sd and se using plyr
library(plyr)
NosemaSummary <- ddply(QueensNosemaDF, c("Origin", "NosemaDay"), summarise, 
                     n = length(Nosema),
                     mean = mean(Nosema),
                     sd = sd(Nosema),
                     se = sd / sqrt(n))

print(NosemaSummary)
```


```{r}

#Create plot in ggplot 
plot <- ggplot(data = NosemaSummary, 
               aes(x = NosemaDay, 
                   y = mean, 
                   group = Origin, 
                   colour = Origin)
) + geom_line(size=1.5) + geom_point(size=4) + scale_colour_manual(values = c("dodgerblue4", "black")) + labs(x = "Time (days after requeening)", y = "Nosmea Load (spores/bee)") + coord_cartesian(ylim = c(0, 5000000), xlim = c(0,60)) + geom_errorbar(aes(ymin = mean - se, ymax = mean + se, width = 0.9)) 

# add a theme and add asterix for significance 
plot + scale_fill_brewer(palette = "Paired") + theme_minimal(base_size = 17) + annotate(geom = "text", x = 34, y = 4500000, label = "*",cex = 8) + annotate(geom = "text", x = 55, y = 2850000, label = "*",cex = 8) 

```

```{r}
# stats comparing treatemetns at each timepoint:

#Check for normality and log transform
#hist(QueensDF$Nosema)
#hist(log(QueensDF$Nosema))

#Subset data into 3 time steps (T1-T4)
NosemaT1 <- QueensDF[c(1:40),]
NosemaT2 <- QueensDF[c(78:106),]
NosemaT3 <- QueensDF[c(107:132),]


# run 3 t-tests comparing california to local colony mass
ComparingQueensNosemaT1 <- aov(Nosema~Origin, data=NosemaT1)
ComparingQueensNosemaT2 <- aov(Nosema~Origin, data=NosemaT2)
ComparingQueensNosemaT3 <- aov(Nosema~Origin, data=NosemaT3)



# checking residuals for T1 for normality 
ComparingQueensNosemaT1.res = resid(ComparingQueensNosemaT1)

NosemaT1 <- NosemaT1[-40,]

#plot(log(NosemaT1$Nosema), ComparingQueensNosemaT1.res) 
#abline(0, 0)          


#store the summaries in some variables
w <- summary(ComparingQueensNosemaT1) #Not sig
x <- summary(ComparingQueensNosemaT2) #Sig
y <- summary(ComparingQueensNosemaT3) #Sig

```

```{r}

# create summary and sd and se using plyr for varroa
# Remove rows containing NAs in column Varroa
VarroaDF <- QueensDF[!is.na(QueensDF$Varroa),]

VarroaSummary <- ddply(VarroaDF, c("Origin", "VarroaDay"), summarise, 
                       n = length(Varroa),
                       mean = mean(Varroa),
                       sd = sd(Varroa),
                       se = sd / sqrt(n))

```

```{r}
#Create varroa line graph in ggplot 
plot <- ggplot(data = VarroaSummary, 
               aes(x = VarroaDay, 
                   y = mean, 
                   group = Origin, 
                   colour = Origin)
) + geom_line(size=1.5) + geom_point(size=4) + scale_colour_manual(values = c("dodgerblue4", "black")) + labs(x = "Time (days after requeening)", y = "Varroa Load (mites/300 bees)") + coord_cartesian(ylim = c(0, 12)) + geom_errorbar(aes(ymin = mean - se, ymax = mean + se, width = .8)) 

# add a theme and add asterix for significance 
plot + scale_fill_brewer(palette = "Paired") + theme_minimal(base_size = 17)  

```

```{r}
# pollen work making a bargraph
# read in data
PollenDF <- read.table("PollenQueens.csv",header=TRUE,sep=",",stringsAsFactors=FALSE)

PollenDF2 <- PollenDF[!is.na(PollenDF$Pollen),]

PollenDF2$WeekChar <- as.character(PollenDF2$Week)

PollenSummary <- ddply(PollenDF2, c("Origin", "WeekChar"), summarise, 
                       n = length(Pollen),
                       mean = mean(Pollen),
                       sd = sd(Pollen),
                       se = sd / sqrt(n))

```


```{r}
colors <- c("dodgerblue4", "slategray3")

plot2 <- ggplot(PollenSummary, aes(x=WeekChar, y=mean, fill=Origin)) + 
  geom_bar(stat="identity", 
           position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=.4,
                position=position_dodge(.9)) + labs(x="Time (weeks after requeening)", 
                                                    y = "Pollen Weight (grams)")

plot2 + theme_minimal(base_size = 17) + coord_cartesian(ylim = c(0, 40)) + scale_fill_manual(values=colors) + annotate(geom = "text", x = 5, y = 27, label = "*",cex = 6) 

```


```{r}
# Brood line graph

BroodDF <- PollenDF[!is.na(PollenDF$Brood),]

BroodSummary <- ddply(BroodDF, c("Origin", "BroodDays"), summarise, 
                       n = length(Brood),
                       mean = mean(Brood),
                       sd = sd(Brood),
                       se = sd / sqrt(n))

```


```{r}
plot <- ggplot(data = BroodSummary, 
               aes(x = BroodDays, 
                   y = mean, 
                   group = Origin, 
                   colour = Origin)
) + geom_line(size=1.5) + geom_point(size=4) + scale_colour_manual(values = c("dodgerblue4", "black")) + labs(x = "Time (days after requeening)", y = "Frames of Brood/Hive") + coord_cartesian(ylim = c(2, 7)) + geom_errorbar(aes(ymin = mean - se, ymax = mean + se, width = 1)) 

# add a theme and add asterix for significance 
plot + scale_fill_brewer(palette = "Paired") + theme_minimal(base_size = 17) + annotate(geom = "text", x = 15, y = 5.2, label = "*",cex = 8) + annotate(geom = "text", x = 43, y = 6, label = "*",cex = 8) + annotate(geom = "text", x = 64, y = 6.7, label = "*",cex = 8) 

```


```{r}
# split data frame and do t-tests between treatments and each time point
x<-split(BroodDF, BroodDF$BroodDays)

t.test(x$`15`$Brood~x$`15`$Origin)
t.test(x$`27`$Brood~x$`27`$Origin)
t.test(x$`43`$Brood~x$`43`$Origin)
t.test(x$`64`$Brood~x$`64`$Origin)

```


```{r}

#Virus stats:

# read in data frame: 
VirusDF <- read.table("RNAVirus.csv",header=TRUE,sep=",",stringsAsFactors=FALSE)

VirusDF <- VirusDF[!is.na(VirusDF$Intensity),]

Virus1 <- split(VirusDF,VirusDF$Virus)
DWV <- Virus1$DWV
IAPV <- Virus1$IAPV

x <- split(DWV, DWV$Time)

t.test(x$T2$Intensity~x$T2$Origin, alternative="greater")


head(x$T2)

x <- x$T2[-c(8),]
x

t.test(x$Intensity~x$Origin, alternative="greater")

```


```{r}

VirusSummary <- ddply(DWV, c("Origin", "Time"), summarise, 
                      n = length(Intensity),
                      mean = mean(Intensity),
                      sd = sd(Intensity),
                      se = sd / sqrt(n))


```

```{r}
#Plot Virus in ggplot

colors <- c("dodgerblue4", "slategray3")

plot1 <- ggplot(VirusSummary, aes(x=Time, y=mean, fill=Origin)) + 
  geom_bar(stat="identity", 
           position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=.4,
                position=position_dodge(.9)) + labs(x="Time (28 days apart)", 
                                                    y = "DWV Volume (Int)")

plot1 + theme_minimal(base_size = 17) + coord_cartesian(ylim = c(0, 4)) + scale_fill_manual(values=colors)


```


```{r}
#IAPV
VirusDF <- VirusDF[!is.na(VirusDF$Intensity),]

Virus1 <- split(VirusDF,VirusDF$Virus)
IAPV <- Virus1$IAPV

z <- split(IAPV, IAPV$Time)

t.test(z$T2$Intensity~z$T2$Origin)
t.test(z$T1$Intensity~z$T1$Origin)

VirusSummary1 <- ddply(IAPV, c("Origin", "Time"), summarise, 
                      n = length(Intensity),
                      mean = mean(Intensity),
                      sd = sd(Intensity),
                      se = sd / sqrt(n))


```

```{r}
#Plot IAPV in ggplot


plot2 <- ggplot(VirusSummary1, aes(x=Time, y=mean, fill=Origin)) + 
  geom_bar(stat="identity", 
           position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=.4,
                position=position_dodge(.9)) + labs(x="Time (28 days apart", 
                                                    y = "IAPV Volume (Int)")

plot2 + theme_minimal(base_size = 17) + coord_cartesian(ylim = c(0, 3)) + scale_fill_manual(values=colors) + annotate(geom = "text", x = 1, y = 2.3, label = "*",cex = 8) + annotate(geom = "text", x = 2, y = 3, label = "*",cex = 8) +scale_x_continuous(breaks = round(seq(min(dat$x), max(dat$x), by = 10),1))




```


```{r}


library(lme4)

glmer(formula = , data = )


#------------------------------------------------------------------------
# MASS

# summary stats for plotting purposes:
MassSummary <- ddply(QueenDF, c("Origin", "MassDay"), summarise, 
                     n = length(Mass),
                     mean = mean(Mass, na.rm = TRUE),
                     sd = sd(Mass, na.rm = TRUE),
                     se = sd / sqrt(n))

# remove time steps with missing data:
MassSummary <- MassSummary[-c(5,10),]
print(MassSummary)



#Create plot in ggplot for mass without x label:
plot <- ggplot(data = MassSummary, 
               aes(x = MassDay, 
                   y = mean, 
                   group = Origin)
) + geom_point(size=3) + scale_colour_manual(values = c("black", "black")) + labs(x = NULL, y = "Colony Mass (lbs)", group="the") + coord_cartesian(ylim = c(20, 80), xlim = c(-5,70)) + geom_errorbar(aes(ymin = mean - se, ymax = mean + se, width = 0.9)) + geom_line(aes(linetype=Origin), size=1) + scale_fill_brewer(palette = "Paired") + theme_minimal(base_size = 17) + theme(legend.position=c(.15, .75), axis.title.y=element_text(margin=margin(0,20,0,0))) + scale_x_continuous(breaks=pretty_breaks(n=10)) + labs(linetype="Queen Origin") + annotate(geom = "text", x = 70, y = 75, label = "A",cex = 6) 

#------------------------------------------------------------------------
# BROOD

# summary stats for plotting purposes:
BroodSummary <- ddply(QueenDF, c("Origin", "BroodDay"), summarise, 
                     n = length(Brood),
                     mean = mean(Brood, na.rm = TRUE),
                     sd = sd(Brood, na.rm = TRUE),
                     se = sd / sqrt(n))
print(BroodSummary)

#Create plot in ggplot for brood 
plot1 <- ggplot(data = BroodSummary, 
               aes(x = BroodDay, 
                   y = mean, 
                   group = Origin)
) + geom_point(size=3) + scale_colour_manual(values = c("black", "black")) + labs(x = "Time (days after requeening)", y = "Frames of Brood") + coord_cartesian(ylim = c(2, 7), xlim = c(-5,70)) + geom_errorbar(aes(ymin = mean - se, ymax = mean + se, width = 0.9)) + geom_line(aes(linetype=Origin), size=1) + scale_fill_brewer(palette = "Paired") + theme_minimal(base_size = 17) + theme(legend.position=c(3, 3), axis.title.y=element_text(margin=margin(0,20,0,0))) + scale_x_continuous(breaks=pretty_breaks(n=10)) + annotate(geom = "text", x = 70, y = 6.5, label = "B",cex = 6) 


# use gridExtra package to combine the two figures:
library(gridExtra)
grid.newpage()
grid.draw(rbind(ggplotGrob(plot), ggplotGrob(plot1), size = "last"))




```

